---
title: "West Boundary Agricultural Water Security"
subtitle: "Pilot dashboard — Kettle River watershed, Midway and upstream (RDKB Electoral Area E)"
format:
  dashboard:
    theme: cosmo
    nav-buttons:
      - icon: info-circle
        href: "#about-data"
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false

library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(plotly)
library(leaflet)
library(tidyhydat)
library(knitr)

# ---- Committed historical baselines (precomputed from HYDAT) -----------------
stations_meta <- read.csv("data/stations.csv", stringsAsFactors = FALSE)
normals       <- read.csv("data/flow_normals.csv", stringsAsFactors = FALSE)
monthly       <- read.csv("data/flow_monthly.csv", stringsAsFactors = FALSE)

hero_id      <- "08NN026"               # Kettle River near Westbridge (main stem)
env_stations <- c("08NN026", "08NN003", "08NN015", "08NN019")  # have real records
ids          <- stations_meta$STATION_NUMBER
yr           <- year(Sys.Date())

# ---- Live readings (current year to date) -----------------------------------
# Wrapped in tryCatch so a temporary ECCC outage can never break the build.
live <- tryCatch(
  realtime_ws(
    station_number = ids,
    parameters     = 47,                # discharge, m3/s
    start_date     = as.Date(paste0(yr, "-01-01")),
    end_date       = Sys.Date()
  ) %>%
    mutate(Date = as.Date(Date)) %>%
    group_by(STATION_NUMBER, Date) %>%
    summarise(flow = mean(Value, na.rm = TRUE), .groups = "drop"),
  error = function(e) {
    message("Live hydrometric fetch failed: ", conditionMessage(e))
    NULL
  }
)
have_live <- !is.null(live) && nrow(live) > 0

# ---- Classify current flow against the seasonal normal range ----------------
classify <- function(flow, p10, p25, p50, p75, p90) {
  dplyr::case_when(
    is.na(flow) | is.na(p50) ~ NA_character_,
    flow <  p10  ~ "Severe Low",
    flow <  p25  ~ "Low",
    flow <= p75  ~ "Normal",
    flow <= p90  ~ "Above Normal",
    TRUE         ~ "High"
  )
}
status_levels <- c("Severe Low", "Low", "Normal", "Above Normal", "High")
status_pal    <- c("Severe Low" = "#d62728", "Low" = "#ff7f0e",
                   "Normal" = "#2ca02c", "Above Normal" = "#1f77b4",
                   "High" = "#17becf")
vb_color      <- c("Severe Low" = "danger", "Low" = "warning", "Normal" = "success",
                   "Above Normal" = "info", "High" = "primary", "No data" = "secondary")

if (have_live) {
  current <- live %>%
    group_by(STATION_NUMBER) %>%
    arrange(Date) %>%
    summarise(last_date = last(Date), flow = last(flow), .groups = "drop") %>%
    mutate(yday = yday(last_date)) %>%
    left_join(normals, by = c("STATION_NUMBER", "yday")) %>%
    mutate(status     = classify(flow, p10, p25, p50, p75, p90),
           pct_median = ifelse(!is.na(p50), 100 * flow / p50, NA_real_)) %>%
    left_join(stations_meta, by = "STATION_NUMBER")
} else {
  current <- stations_meta %>%
    mutate(last_date = as.Date(NA), flow = NA_real_, yday = NA_integer_,
           p10 = NA, p25 = NA, p50 = NA, p75 = NA, p90 = NA,
           status = NA_character_, pct_median = NA_real_)
}

# ---- Groundwater: B.C. observation wells (PGOWN) ----------------------------
gw_wells   <- read.csv("data/gw_wells.csv", stringsAsFactors = FALSE)
gw_normals <- read.csv("data/gw_normals.csv", stringsAsFactors = FALSE)
gw_base    <- "https://www.env.gov.bc.ca/wsd/data_searches/obswell/map/data/"

read_gw_recent <- function(id) {
  tryCatch(
    read_csv(paste0(gw_base, id, "-recent.csv"), show_col_types = FALSE,
             col_types = cols(Time = col_character(), Value = col_double(),
                              .default = col_character())) %>%
      transmute(well = id, Date = as.Date(ymd_hm(Time, quiet = TRUE)), depth = Value) %>%
      filter(!is.na(Date), !is.na(depth)) %>%
      group_by(well, Date) %>% summarise(depth = mean(depth), .groups = "drop"),
    error = function(e) NULL
  )
}
gw_live <- tryCatch(bind_rows(lapply(gw_wells$well, read_gw_recent)),
                    error = function(e) NULL)
have_gw <- !is.null(gw_live) && nrow(gw_live) > 0

# For groundwater a LARGER depth-to-water means a LOWER water table (drier),
# so the percentile comparison is inverted relative to streamflow.
gw_classify <- function(depth, p10, p25, p50, p75, p90) {
  dplyr::case_when(
    is.na(depth) | is.na(p50) ~ NA_character_,
    depth >  p90 ~ "Severe Low",
    depth >  p75 ~ "Low",
    depth >= p25 ~ "Normal",
    depth >= p10 ~ "Above Normal",
    TRUE         ~ "High"
  )
}

if (have_gw) {
  gw_current <- gw_live %>%
    group_by(well) %>% arrange(Date) %>%
    summarise(last_date = last(Date), depth = last(depth), .groups = "drop") %>%
    mutate(yday = yday(last_date)) %>%
    left_join(gw_normals, by = c("well", "yday")) %>%
    mutate(status = gw_classify(depth, p10, p25, p50, p75, p90)) %>%
    left_join(gw_wells, by = "well")
} else {
  gw_current <- gw_wells %>%
    mutate(last_date = as.Date(NA), depth = NA_real_, yday = NA_integer_,
           p10 = NA, p25 = NA, p50 = NA, p75 = NA, p90 = NA, status = NA_character_)
}
gw_below <- sum(gw_current$status %in% c("Low", "Severe Low"), na.rm = TRUE)
gw_n     <- sum(!is.na(gw_current$status))

gw_box <- function(id) {
  w <- gw_current %>% filter(well == id)
  st <- if (nrow(w) && !is.na(w$status)) w$status else "No data"
  list(icon = "moisture", color = vb_color[[st]],
       value = if (!nrow(w) || is.na(w$depth)) "n/a" else paste0(round(w$depth, 2), " m"))
}

h <- current %>% filter(STATION_NUMBER == hero_id)
hflow   <- if (nrow(h)) round(h$flow, 1)        else NA
hpct    <- if (nrow(h)) round(h$pct_median)     else NA
hstatus <- if (nrow(h) && !is.na(h$status)) h$status else "No data"
hdate   <- if (nrow(h)) h$last_date             else NA

n_below <- sum(current$status %in% c("Low", "Severe Low"), na.rm = TRUE)
n_env   <- sum(!is.na(current$status))
irrigation_active <- month(Sys.Date()) %in% 4:10
last_updated <- if (have_live) format(max(live$Date), "%b %d, %Y") else "unavailable"
```

# Current Conditions

## Row {height="22%"}

```{r}
#| content: valuebox
#| title: "Kettle R. near Westbridge — flow"
list(
  icon  = "droplet-half",
  color = vb_color[[hstatus]],
  value = if (is.na(hflow)) "n/a" else paste0(hflow, " m³/s")
)
```

```{r}
#| content: valuebox
#| title: "Percent of seasonal median"
list(
  icon  = "graph-down-arrow",
  color = vb_color[[hstatus]],
  value = if (is.na(hpct)) "n/a" else paste0(hpct, "%")
)
```

```{r}
#| content: valuebox
#| title: "Current flow status"
list(
  icon  = "exclamation-triangle",
  color = vb_color[[hstatus]],
  value = hstatus
)
```

```{r}
#| content: valuebox
#| title: "Monitored sites below normal"
list(
  icon  = "reception-2",
  color = if (n_below > 0) "warning" else "success",
  value = if (n_env == 0) "n/a" else paste0(n_below, " of ", n_env)
)
```

```{r}
#| content: valuebox
#| title: "Irrigation season"
list(
  icon  = "moisture",
  color = if (irrigation_active) "primary" else "secondary",
  value = if (irrigation_active) "Active" else "Off-season"
)
```

## Row {height="78%"}

### Column {width="62%"}

```{r}
#| title: "This year's flow vs. the historical normal range — Kettle River near Westbridge"

env <- normals %>%
  filter(STATION_NUMBER == hero_id, yday <= 365) %>%
  arrange(yday) %>%
  mutate(date = as.Date(yday - 1, origin = paste0(yr, "-01-01")))

cur_line <- if (have_live) {
  live %>% filter(STATION_NUMBER == hero_id) %>% arrange(Date)
} else {
  data.frame(Date = as.Date(character()), flow = numeric())
}

p <- plot_ly() %>%
  add_ribbons(data = env, x = ~date, ymin = ~p10, ymax = ~p90,
              name = "Historical range (10–90%)", line = list(width = 0),
              fillcolor = "rgba(31,119,180,0.12)", hoverinfo = "skip") %>%
  add_ribbons(data = env, x = ~date, ymin = ~p25, ymax = ~p75,
              name = "Typical range (25–75%)", line = list(width = 0),
              fillcolor = "rgba(31,119,180,0.28)", hoverinfo = "skip") %>%
  add_lines(data = env, x = ~date, y = ~p50, name = "Median (normal)",
            line = list(color = "#1f77b4", dash = "dash", width = 2)) %>%
  add_lines(data = cur_line, x = ~Date, y = ~flow, name = paste(yr, "observed"),
            line = list(color = "#d62728", width = 3)) %>%
  layout(
    yaxis = list(title = "Daily mean discharge (m³/s)"),
    xaxis = list(title = "", tickformat = "%b"),
    hovermode = "x unified",
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.08)
  )
p
```

### Column {width="38%"}

```{r}
#| title: "Project-area monitoring network"

communities <- data.frame(
  name = c("Midway", "Rock Creek", "Beaverdell", "Bridesville", "Westbridge"),
  lat  = c(49.0093, 49.0625, 49.4253, 49.0331, 49.1700),
  lng  = c(-118.7726, -118.9214, -119.0828, -119.1564, -118.9600)
)

mdat <- current %>%
  mutate(col = ifelse(is.na(status), "#7f7f7f", status_pal[status]),
         lab = ifelse(is.na(flow), "no data",
                      paste0(round(flow, 1), " m³/s")))

leaflet(mdat) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addLabelOnlyMarkers(
    data = communities, ~lng, ~lat, label = ~name,
    labelOptions = labelOptions(noHide = TRUE, direction = "top", textOnly = TRUE,
      style = list("font-weight" = "600", "color" = "#444", "font-size" = "12px"))
  ) %>%
  addCircleMarkers(
    ~LONGITUDE, ~LATITUDE, radius = 10, color = ~col, weight = 2,
    stroke = TRUE, fillOpacity = 0.85,
    label = ~lapply(paste0("<b>", STATION_NAME, "</b>"), htmltools::HTML),
    popup = ~paste0(
      "<b>", STATION_NAME, "</b><br>", STATION_NUMBER,
      "<br>Flow: ", ifelse(is.na(flow), "n/a", paste0(round(flow, 1), " m³/s")),
      "<br>% of median: ", ifelse(is.na(pct_median), "n/a", paste0(round(pct_median), "%")),
      "<br>Status: ", ifelse(is.na(status), "no baseline", status)
    )
  ) %>%
  addLegend("bottomright", colors = unname(status_pal),
            labels = names(status_pal), title = "Flow status", opacity = 0.9)
```

```{r}
#| title: "Current status by station"

current %>%
  arrange(pct_median) %>%
  transmute(
    Station = STATION_NAME,
    `Flow (m³/s)` = round(flow, 1),
    `% of median` = round(pct_median),
    Status = ifelse(is.na(status), "no baseline", status)
  ) %>%
  kable()
```

# Seasonal & Drought Risk

## Row

### Column {width="55%"}

```{r}
#| title: "Monthly flow climatology — Kettle River near Westbridge"

mh <- monthly %>% filter(STATION_NUMBER == hero_id) %>%
  mutate(month = factor(month.abb[month], levels = month.abb))

cur_month <- if (have_live) {
  live %>% filter(STATION_NUMBER == hero_id) %>%
    mutate(month = factor(month.abb[month(Date)], levels = month.abb)) %>%
    group_by(month) %>% summarise(m = mean(flow), .groups = "drop")
} else data.frame(month = factor(character(), levels = month.abb), m = numeric())

plot_ly() %>%
  add_boxplot(data = mh, x = ~month, y = ~monthly_mean, name = "Historical (per year)",
              boxpoints = FALSE, line = list(color = "#1f77b4"),
              fillcolor = "rgba(31,119,180,0.25)") %>%
  add_markers(data = cur_month, x = ~month, y = ~m, name = paste(yr),
              marker = list(color = "#d62728", size = 11, symbol = "diamond")) %>%
  layout(yaxis = list(title = "Monthly mean discharge (m³/s)"),
         xaxis = list(title = ""),
         legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.1))
```

### Column {width="45%"}

```{r}
#| title: "How does each site compare to its seasonal normal?"

bar <- current %>% filter(!is.na(pct_median)) %>% arrange(pct_median)

if (nrow(bar) > 0) {
  plot_ly(bar, x = ~pct_median, y = ~reorder(STATION_NAME, pct_median),
          type = "bar", orientation = "h",
          marker = list(color = status_pal[bar$status]),
          hovertemplate = "%{y}: %{x:.0f}% of median<extra></extra>") %>%
    layout(
      xaxis = list(title = "Current flow as % of seasonal median"),
      yaxis = list(title = ""),
      shapes = list(list(type = "line", x0 = 100, x1 = 100, y0 = -0.5,
                         y1 = nrow(bar) - 0.5, line = list(color = "#444", dash = "dot")))
    )
} else {
  cat("Current comparison unavailable — live data could not be retrieved.")
}
```

## Row {height="28%"}

::: {.card title="Why this matters for West Boundary agriculture"}
Summer is when agricultural water demand peaks and natural supply is lowest. The
West Boundary has faced repeated drought stress, including the **2021 Temporary
Protection Order** in the West Kettle watershed that restricted withdrawals.

The envelope chart and the comparisons above put **today's flows in the context
of decades of record** (48–70 years per main station), so producers and planners
can quickly see whether the watershed is tracking toward a low-flow year. When
the red line sits near or below the lower edge of the blue band, the watershed is
running drier than usual for the season — an early signal for storage,
conservation, and engagement decisions.
:::

# Groundwater & Aquifers

## Row {height="22%"}

```{r}
#| content: valuebox
#| title: "Grand Forks well (OW217) — depth to water"
gw_box("OW217")
```

```{r}
#| content: valuebox
#| title: "Beaverdell well (OW306) — depth to water"
gw_box("OW306")
```

```{r}
#| content: valuebox
#| title: "Midway well (OW444) — depth to water"
gw_box("OW444")
```

```{r}
#| content: valuebox
#| title: "Observation wells below normal"
list(
  icon  = "reception-2",
  color = if (gw_below > 0) "warning" else "success",
  value = if (gw_n == 0) "n/a" else paste0(gw_below, " of ", gw_n)
)
```

## Row {height="78%"}

### Column {width="62%"}

```{r}
#| title: "Water-table depth vs. the historical normal range — Grand Forks (OW217). Lower on the chart = deeper water table = drier."

gw_hero <- "OW217"
genv <- gw_normals %>%
  filter(well == gw_hero, yday <= 365) %>%
  arrange(yday) %>%
  mutate(date = as.Date(yday - 1, origin = paste0(yr, "-01-01")))

gcur <- if (have_gw) {
  gw_live %>%
    filter(well == gw_hero, Date >= as.Date(paste0(yr, "-01-01"))) %>%
    arrange(Date)
} else {
  data.frame(Date = as.Date(character()), depth = numeric())
}

plot_ly() %>%
  add_ribbons(data = genv, x = ~date, ymin = ~p10, ymax = ~p90,
              name = "Historical range (10–90%)", line = list(width = 0),
              fillcolor = "rgba(44,160,44,0.12)", hoverinfo = "skip") %>%
  add_ribbons(data = genv, x = ~date, ymin = ~p25, ymax = ~p75,
              name = "Typical range (25–75%)", line = list(width = 0),
              fillcolor = "rgba(44,160,44,0.28)", hoverinfo = "skip") %>%
  add_lines(data = genv, x = ~date, y = ~p50, name = "Median (normal)",
            line = list(color = "#2ca02c", dash = "dash", width = 2)) %>%
  add_lines(data = gcur, x = ~Date, y = ~depth, name = paste(yr, "observed"),
            line = list(color = "#8c564b", width = 3)) %>%
  layout(
    yaxis = list(title = "Depth to water below ground (m)", autorange = "reversed"),
    xaxis = list(title = "", tickformat = "%b"),
    hovermode = "x unified",
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.08)
  )
```

### Column {width="38%"}

```{r}
#| title: "Observation wells (B.C. PGOWN)"

gw_current %>%
  transmute(
    Well = well,
    Location = location,
    Aquifer = aquifer,
    `Depth (m)` = round(depth, 2),
    Record = ifelse(is.na(start_year), "—", paste0(start_year, "–", end_year)),
    Status = ifelse(is.na(status), "no baseline", status)
  ) %>%
  kable()
```

::: {.card title="Why groundwater matters here"}
The Kettle valley aquifers are **strongly connected to the river** — the Grand
Forks aquifer sits next to the Kettle/Granby confluence and rises and falls with
the freshet. When both streamflow **and** water tables are running below normal at
the same time, it points to a basin-wide moisture deficit rather than a local
blip — an important signal for irrigation timing, storage, and shared
conservation decisions.
:::

# Stations & Live Data

## Row {height="40%"}

```{r}
#| title: "Monitoring stations (Water Survey of Canada)"

stations_meta %>%
  arrange(STATION_NUMBER) %>%
  transmute(
    `Station ID` = STATION_NUMBER,
    Name = STATION_NAME,
    `Drainage area (km²)` = round(DRAINAGE_AREA_GROSS),
    `Record` = ifelse(is.na(start_year), "—", paste0(start_year, "–", end_year)),
    `Years` = n_years
  ) %>%
  kable()
```

## Row {height="60%"}

```{r}
#| title: "Daily mean discharge — all monitored sites, current year"

if (have_live) {
  lab <- setNames(stations_meta$STATION_NAME, stations_meta$STATION_NUMBER)
  live2 <- live %>% mutate(name = lab[STATION_NUMBER])
  plot_ly(live2, x = ~Date, y = ~flow, color = ~name, type = "scatter",
          mode = "lines") %>%
    layout(yaxis = list(title = "Discharge (m³/s)"), xaxis = list(title = ""),
           hovermode = "x unified",
           legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.12))
} else {
  cat("Live data could not be retrieved at build time.")
}
```

# About & Data {#about-data}

## Row

### Column

::: {.card title="What this dashboard is"}
A **pilot dashboard** for the RDKB West Boundary Agricultural Water Supply and
Storage planning process. It brings public hydrometric information together in an
accessible place to support **planning and community engagement** — helping
producers, First Nations, local government, and agencies share a common picture
of seasonal water conditions and drought risk in the Kettle River watershed from
Midway upstream.

It is **not** a regulatory monitoring tool. Readings are provisional and
unverified, and should not be used for compliance or enforcement.
:::

::: {.card title="Data sources & methods"}
- **Live streamflow:** Environment and Climate Change Canada (ECCC) real-time
  hydrometric web service, retrieved with the open-source
  [`tidyhydat`](https://docs.ropensci.org/tidyhydat/) R package (parameter 47,
  daily mean, m³/s).
- **Streamflow normals:** Water Survey of Canada **HYDAT** database. For each day
  of the year we pool all daily flows within a ±7-day window across the full
  record and compute the 10th–90th percentile envelope and median. Baselines are
  precomputed by `scripts/build_flow_normals.R` and stored as small CSVs in
  `/data`, so the published site never needs the full 1.3 GB database.
- **Groundwater:** B.C. Provincial Groundwater Observation Well Network (PGOWN),
  Open Government Licence – B.C. Live readings and historical depth-to-water
  baselines use the same percentile method (`scripts/build_gw_normals.R`).
- **Status classes:** *Severe Low* (least water available — below the 10th
  percentile for the day), then *Low*, *Normal*, *Above Normal*, *High*.
:::

### Column

::: {.card title="Confidentiality & responsible use"}
This dashboard uses **public and voluntarily shared information only**. Should
producer-contributed inputs (for example well levels, repeat photo points, or
irrigation timing) be added in future, they would be **aggregated or anonymised**
so that individual operations cannot be identified, consistent with the project's
commitment to a transparent, non-regulatory process and with FOIPPA.
:::

::: {.card title="Ownership, hosting & upkeep"}
Built entirely with **open-source tools** (Quarto + R) so it can be fully owned,
hosted, and maintained by the Regional District. All source files, data-build
scripts, and documentation are included in the project repository. Refreshing the
historical baselines is a single command (`Rscript scripts/build_flow_normals.R`);
live readings update automatically each time the site is rebuilt.

*Potential future inputs identified in the RFP — groundwater well levels,
date-stamped/repeat photo points, and irrigation season timing — can be added as
additional pages using the same pattern.*
:::
