---
title: "West Boundary Agricultural Water Security"
subtitle: "Pilot dashboard — Kettle River watershed, Midway and upstream (RDKB Electoral Area E)"
format:
  dashboard:
    theme: cosmo
    nav-buttons:
      - icon: info-circle
        href: "#about-data"
    css: dashboard.css
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false

library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(plotly)
library(leaflet)
library(tidyhydat)
library(knitr)

# ---- Committed historical baselines (precomputed from HYDAT) -----------------
stations_meta <- read.csv("data/stations.csv", stringsAsFactors = FALSE)
normals       <- read.csv("data/flow_normals.csv", stringsAsFactors = FALSE)
monthly       <- read.csv("data/flow_monthly.csv", stringsAsFactors = FALSE)
flow_mad      <- read.csv("data/flow_mad.csv", stringsAsFactors = FALSE)
flow_analog   <- read.csv("data/flow_analog_years.csv", stringsAsFactors = FALSE)
flow_trend    <- read.csv("data/flow_trend.csv", stringsAsFactors = FALSE)
flow_summer   <- read.csv("data/flow_summer_low.csv", stringsAsFactors = FALSE)
snow_station  <- read.csv("data/snow_station.csv", stringsAsFactors = FALSE)
snow_normals  <- read.csv("data/snow_normals.csv", stringsAsFactors = FALSE)
snow_cur_wy   <- read.csv("data/snow_current_wy.csv", stringsAsFactors = FALSE) %>%
  mutate(Date = as.Date(Date))

hero_id      <- "08NN026"               # Kettle River near Westbridge (main stem)
env_stations <- c("08NN026", "08NN003", "08NN015", "08NN019")  # have real records
ids          <- stations_meta$STATION_NUMBER
yr           <- year(Sys.Date())

# ---- Live readings (current year to date) -----------------------------------
# Wrapped in tryCatch so a temporary ECCC outage can never break the build.
live <- tryCatch(
  realtime_ws(
    station_number = ids,
    parameters     = 47,                # discharge, m3/s
    start_date     = as.Date(paste0(yr, "-01-01")),
    end_date       = Sys.Date()
  ) %>%
    mutate(Date = as.Date(Date)) %>%
    group_by(STATION_NUMBER, Date) %>%
    summarise(flow = mean(Value, na.rm = TRUE), .groups = "drop"),
  error = function(e) {
    message("Live hydrometric fetch failed: ", conditionMessage(e))
    NULL
  }
)
have_live <- !is.null(live) && nrow(live) > 0

# ---- Classify current flow against the seasonal normal range ----------------
classify <- function(flow, p10, p25, p50, p75, p90) {
  dplyr::case_when(
    is.na(flow) | is.na(p50) ~ NA_character_,
    flow <  p10  ~ "Severe Low",
    flow <  p25  ~ "Low",
    flow <= p75  ~ "Normal",
    flow <= p90  ~ "Above Normal",
    TRUE         ~ "High"
  )
}
status_levels <- c("Severe Low", "Low", "Normal", "Above Normal", "High")
status_pal    <- c("Severe Low" = "#d62728", "Low" = "#ff7f0e",
                   "Normal" = "#2ca02c", "Above Normal" = "#1f77b4",
                   "High" = "#17becf")
vb_color      <- c("Severe Low" = "danger", "Low" = "warning", "Normal" = "success",
                   "Above Normal" = "info", "High" = "primary", "No data" = "secondary")

if (have_live) {
  current <- live %>%
    group_by(STATION_NUMBER) %>%
    arrange(Date) %>%
    summarise(last_date = last(Date), flow = last(flow), .groups = "drop") %>%
    mutate(yday = yday(last_date)) %>%
    left_join(normals, by = c("STATION_NUMBER", "yday")) %>%
    mutate(status     = classify(flow, p10, p25, p50, p75, p90),
           pct_median = ifelse(!is.na(p50), 100 * flow / p50, NA_real_)) %>%
    left_join(stations_meta, by = "STATION_NUMBER")
} else {
  current <- stations_meta %>%
    mutate(last_date = as.Date(NA), flow = NA_real_, yday = NA_integer_,
           p10 = NA, p25 = NA, p50 = NA, p75 = NA, p90 = NA,
           status = NA_character_, pct_median = NA_real_)
}

# ---- Groundwater: B.C. observation wells (PGOWN) ----------------------------
gw_wells   <- read.csv("data/gw_wells.csv", stringsAsFactors = FALSE)
gw_normals <- read.csv("data/gw_normals.csv", stringsAsFactors = FALSE)
gw_base    <- "https://www.env.gov.bc.ca/wsd/data_searches/obswell/map/data/"

read_gw_recent <- function(id) {
  tryCatch(
    read_csv(paste0(gw_base, id, "-recent.csv"), show_col_types = FALSE,
             col_types = cols(Time = col_character(), Value = col_double(),
                              .default = col_character())) %>%
      transmute(well = id, Date = as.Date(ymd_hm(Time, quiet = TRUE)), depth = Value) %>%
      filter(!is.na(Date), !is.na(depth)) %>%
      group_by(well, Date) %>% summarise(depth = mean(depth), .groups = "drop"),
    error = function(e) NULL
  )
}
gw_live <- tryCatch(bind_rows(lapply(gw_wells$well, read_gw_recent)),
                    error = function(e) NULL)
have_gw <- !is.null(gw_live) && nrow(gw_live) > 0

# For groundwater a LARGER depth-to-water means a LOWER water table (drier),
# so the percentile comparison is inverted relative to streamflow.
gw_classify <- function(depth, p10, p25, p50, p75, p90) {
  dplyr::case_when(
    is.na(depth) | is.na(p50) ~ NA_character_,
    depth >  p90 ~ "Severe Low",
    depth >  p75 ~ "Low",
    depth >= p25 ~ "Normal",
    depth >= p10 ~ "Above Normal",
    TRUE         ~ "High"
  )
}

if (have_gw) {
  gw_current <- gw_live %>%
    group_by(well) %>% arrange(Date) %>%
    summarise(last_date = last(Date), depth = last(depth), .groups = "drop") %>%
    mutate(yday = yday(last_date)) %>%
    left_join(gw_normals, by = c("well", "yday")) %>%
    mutate(status = gw_classify(depth, p10, p25, p50, p75, p90)) %>%
    left_join(gw_wells, by = "well")
} else {
  gw_current <- gw_wells %>%
    mutate(last_date = as.Date(NA), depth = NA_real_, yday = NA_integer_,
           p10 = NA, p25 = NA, p50 = NA, p75 = NA, p90 = NA, status = NA_character_)
}
gw_below <- sum(gw_current$status %in% c("Low", "Severe Low"), na.rm = TRUE)
gw_n     <- sum(!is.na(gw_current$status))

gw_box <- function(id) {
  w <- gw_current %>% filter(well == id)
  st <- if (nrow(w) && !is.na(w$status)) w$status else "No data"
  list(icon = "moisture", color = vb_color[[st]],
       value = if (!nrow(w) || is.na(w$depth)) "n/a" else paste0(round(w$depth, 2), " m"))
}

h <- current %>% filter(STATION_NUMBER == hero_id)
hflow   <- if (nrow(h)) round(h$flow, 1)        else NA
hpct    <- if (nrow(h)) round(h$pct_median)     else NA
hstatus <- if (nrow(h) && !is.na(h$status)) h$status else "No data"
hdate   <- if (nrow(h)) h$last_date             else NA

mad_row   <- flow_mad %>% filter(STATION_NUMBER == hero_id)
mad_val   <- if (nrow(mad_row)) mad_row$mad[1] else NA_real_
thr_20    <- if (nrow(mad_row)) mad_row$threshold_20pct[1] else NA_real_
thr_05    <- if (nrow(mad_row)) mad_row$threshold_05pct[1] else NA_real_
pct_mad   <- if (!is.na(hflow) && !is.na(mad_val)) round(100 * hflow / mad_val) else NA_integer_
mad_alert <- dplyr::case_when(
  is.na(pct_mad)           ~ "No data",
  pct_mad < 5              ~ "Below 5% MAD",
  pct_mad < 20             ~ "Below 20% MAD",
  TRUE                     ~ "Above 20% MAD"
)
mad_vb <- dplyr::case_when(
  mad_alert == "Below 5% MAD"  ~ "danger",
  mad_alert == "Below 20% MAD" ~ "warning",
  mad_alert == "Above 20% MAD" ~ "success",
  TRUE                         ~ "secondary"
)

if (have_live) {
  cum_df <- live %>%
    filter(STATION_NUMBER == hero_id) %>%
    mutate(yday = yday(Date)) %>%
    left_join(normals %>% filter(STATION_NUMBER == hero_id), by = "yday") %>%
    filter(!is.na(flow), !is.na(p50))
  cum_deficit_pct <- if (nrow(cum_df)) {
    round(100 * (1 - sum(cum_df$flow) / sum(cum_df$p50)), 0)
  } else NA_integer_
} else {
  cum_deficit_pct <- NA_integer_
}

trend_row <- flow_trend %>% filter(STATION_NUMBER == hero_id)

# Snowpack (Grano Creek 2E07P) — leading indicator for summer supply
snow_wy <- ifelse(month(Sys.Date()) >= 10, yr + 1L, yr)
snow_line <- snow_cur_wy %>% arrange(Date)
snow_last <- if (nrow(snow_line)) snow_line[nrow(snow_line), , drop = FALSE] else NULL
snow_swe  <- if (!is.null(snow_last)) round(snow_last$swe, 0) else NA_real_
snow_wyday <- if (!is.null(snow_last)) snow_last$wyday[1] else NA_integer_
snow_norm_day <- if (!is.na(snow_wyday)) {
  snow_normals %>% filter(wyday == snow_wyday)
} else data.frame()
snow_pct <- if (nrow(snow_norm_day) && !is.na(snow_swe)) {
  round(100 * snow_swe / snow_norm_day$p50[1])
} else NA_integer_
snow_status <- if (is.na(snow_pct)) "No data" else if (snow_pct < 50) "Well below normal" else if (snow_pct < 80) "Below normal" else "Near normal"
snow_vb <- if (is.na(snow_pct)) "secondary" else if (snow_pct < 50) "danger" else if (snow_pct < 80) "warning" else "success"

n_below <- sum(current$status %in% c("Low", "Severe Low"), na.rm = TRUE)
n_env   <- sum(!is.na(current$status))
irrigation_active <- month(Sys.Date()) %in% 4:10
last_updated <- if (have_live) format(max(live$Date), "%b %d, %Y") else "unavailable"
```

# Current Conditions

## Row {height="22%"}

```{r}
#| content: valuebox
#| title: "Kettle R. near Westbridge — flow"
list(
  icon  = "droplet-half",
  color = vb_color[[hstatus]],
  value = if (is.na(hflow)) "n/a" else paste0(hflow, " m³/s")
)
```

```{r}
#| content: valuebox
#| title: "Percent of Mean Annual Discharge"
list(
  icon  = "speedometer",
  color = mad_vb,
  value = if (is.na(pct_mad)) "n/a" else paste0(pct_mad, "% MAD")
)
```

```{r}
#| content: valuebox
#| title: "Year-to-date volume deficit"
list(
  icon  = "graph-down",
  color = if (!is.na(cum_deficit_pct) && cum_deficit_pct > 15) "danger" else "warning",
  value = if (is.na(cum_deficit_pct)) "n/a" else paste0(cum_deficit_pct, "% below normal")
)
```

```{r}
#| content: valuebox
#| title: "MAD conservation trigger"
list(
  icon  = "sign-stop",
  color = mad_vb,
  value = mad_alert
)
```

```{r}
#| content: valuebox
#| title: "Monitored sites below normal"
list(
  icon  = "reception-2",
  color = if (n_below > 0) "warning" else "success",
  value = if (n_env == 0) "n/a" else paste0(n_below, " of ", n_env)
)
```

## Row {height="78%"}

### Column {width="62%"}

```{r}
#| title: "This year's flow vs. the historical normal range — Kettle River near Westbridge"

env <- normals %>%
  filter(STATION_NUMBER == hero_id, yday <= 365) %>%
  arrange(yday) %>%
  mutate(date = as.Date(yday - 1, origin = paste0(yr, "-01-01")))

cur_line <- if (have_live) {
  live %>% filter(STATION_NUMBER == hero_id) %>% arrange(Date)
} else {
  data.frame(Date = as.Date(character()), flow = numeric())
}

analog_plot <- flow_analog %>%
  mutate(plot_date = as.Date(yday - 1, origin = paste0(yr, "-01-01")))

p <- plot_ly() %>%
  add_ribbons(data = env, x = ~date, ymin = ~p10, ymax = ~p90,
              name = "Historical range (10–90%)", line = list(width = 0),
              fillcolor = "rgba(31,119,180,0.12)", hoverinfo = "skip") %>%
  add_ribbons(data = env, x = ~date, ymin = ~p25, ymax = ~p75,
              name = "Typical range (25–75%)", line = list(width = 0),
              fillcolor = "rgba(31,119,180,0.28)", hoverinfo = "skip") %>%
  add_lines(data = env, x = ~date, y = ~p50, name = "Median (normal)",
            line = list(color = "#1f77b4", dash = "dash", width = 2))

for (ay in sort(unique(analog_plot$analog_year))) {
  ad <- analog_plot %>% filter(analog_year == ay)
  p <- p %>% add_lines(
    data = ad, x = ~plot_date, y = ~flow,
    name = paste(ay, "drought year"),
    line = list(width = 1.5, dash = "dash"),
    visible = "legendonly"
  )
}

p %>%
  add_lines(data = cur_line, x = ~Date, y = ~flow, name = paste(yr, "observed"),
            line = list(color = "#d62728", width = 3)) %>%
  layout(
    yaxis = list(title = "Daily mean discharge (m³/s)"),
    xaxis = list(title = "", tickformat = "%b"),
    hovermode = "x unified",
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.08)
  )
```

### Column {width="38%"}

```{r}
#| title: "Project-area monitoring network"

communities <- data.frame(
  name = c("Midway", "Rock Creek", "Beaverdell", "Bridesville", "Westbridge"),
  lat  = c(49.0093, 49.0625, 49.4253, 49.0331, 49.1700),
  lng  = c(-118.7726, -118.9214, -119.0828, -119.1564, -118.9600)
)

mdat <- current %>%
  mutate(col = ifelse(is.na(status), "#7f7f7f", status_pal[status]),
         lab = ifelse(is.na(flow), "no data",
                      paste0(round(flow, 1), " m³/s")))

leaflet(mdat) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addLabelOnlyMarkers(
    data = communities, ~lng, ~lat, label = ~name,
    labelOptions = labelOptions(noHide = TRUE, direction = "top", textOnly = TRUE,
      style = list("font-weight" = "600", "color" = "#444", "font-size" = "12px"))
  ) %>%
  addCircleMarkers(
    ~LONGITUDE, ~LATITUDE, radius = 10, color = ~col, weight = 2,
    stroke = TRUE, fillOpacity = 0.85,
    label = ~lapply(paste0("<b>", STATION_NAME, "</b>"), htmltools::HTML),
    popup = ~paste0(
      "<b>", STATION_NAME, "</b><br>", STATION_NUMBER,
      "<br>Flow: ", ifelse(is.na(flow), "n/a", paste0(round(flow, 1), " m³/s")),
      "<br>% of median: ", ifelse(is.na(pct_median), "n/a", paste0(round(pct_median), "%")),
      "<br>Status: ", ifelse(is.na(status), "no baseline", status)
    )
  ) %>%
  addLegend("bottomright", colors = unname(status_pal),
            labels = names(status_pal), title = "Flow status", opacity = 0.9)
```

```{r}
#| title: "Current status by station"

current %>%
  arrange(pct_median) %>%
  transmute(
    Station = STATION_NAME,
    `Flow (m³/s)` = round(flow, 1),
    `% of median` = round(pct_median),
    `% MAD` = ifelse(STATION_NUMBER == hero_id, pct_mad, NA_integer_),
    Status = ifelse(is.na(status), "no baseline", status)
  ) %>%
  kable()
```

# Seasonal & Drought Risk

## Row {height="48%"}

### Column {width="58%"}

```{r}
#| title: "Monthly flow climatology — Kettle River near Westbridge"

library(ggplot2)

mh <- monthly %>%
  filter(STATION_NUMBER == hero_id) %>%
  mutate(month_num = month)

monthly_env <- mh %>%
  group_by(month_num) %>%
  summarise(
    p25 = quantile(monthly_mean, 0.25, na.rm = TRUE),
    p50 = quantile(monthly_mean, 0.50, na.rm = TRUE),
    p75 = quantile(monthly_mean, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

cur_month <- if (have_live) {
  live %>%
    filter(STATION_NUMBER == hero_id) %>%
    mutate(month_num = month(Date)) %>%
    group_by(month_num) %>%
    summarise(m = mean(flow), .groups = "drop") %>%
    filter(!is.na(m))
} else {
  data.frame(month_num = integer(), m = numeric())
}

p_month <- ggplot(monthly_env, aes(x = month_num)) +
  geom_ribbon(aes(ymin = p25, ymax = p75), fill = "#1f77b4", alpha = 0.25) +
  geom_line(aes(y = p50), color = "#1f77b4", linewidth = 0.9, linetype = "dashed") +
  geom_point(
    data = cur_month, aes(y = m),
    color = "#d62728", size = 3, shape = 18
  ) +
  scale_x_continuous(breaks = 1:12, labels = month.abb, expand = expansion(mult = 0.02)) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.08))) +
  labs(x = NULL, y = "Monthly mean discharge (m³/s)") +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "none",
    plot.margin = margin(4, 8, 4, 4)
  )

ggplotly(p_month, width = NULL, height = 280, tooltip = "none") %>%
  layout(
    autosize = FALSE,
    height = 280,
    margin = list(l = 50, r = 10, t = 28, b = 32),
    annotations = list(
      list(x = 0, y = 1.12, xref = "paper", yref = "paper", xanchor = "left",
           text = "Shaded: 25–75%  |  Dashed: median  |  ◆: current year",
           showarrow = FALSE, font = list(size = 9, color = "#555"))
    )
  ) %>%
  config(displayModeBar = FALSE, responsive = FALSE)
```

### Column {width="42%"}

```{r}
#| title: "How does each site compare to its seasonal normal?"

bar <- current %>%
  filter(!is.na(pct_median)) %>%
  mutate(
  label = gsub(" NEAR ", "<br>NEAR ", STATION_NAME),
  label = gsub(" AT ", "<br>AT ", label)
) %>%
  arrange(pct_median)

if (nrow(bar) > 0) {
  plot_ly(bar, x = ~pct_median, y = ~reorder(label, pct_median),
          type = "bar", orientation = "h",
          marker = list(color = status_pal[bar$status]),
          hovertemplate = "%{y}: %{x:.0f}% of median<extra></extra>") %>%
    layout(
      autosize = FALSE,
      height = 280,
      margin = list(l = 10, r = 15, t = 10, b = 45),
      xaxis = list(title = "Current flow (% of seasonal median)", automargin = TRUE,
                   titlefont = list(size = 10)),
      yaxis = list(title = "", automargin = TRUE, tickfont = list(size = 9)),
      shapes = list(list(type = "line", x0 = 100, x1 = 100, y0 = -0.5,
                         y1 = nrow(bar) - 0.5, line = list(color = "#444", dash = "dot")))
    ) %>%
    config(displayModeBar = FALSE, responsive = FALSE)
} else {
  cat("Current comparison unavailable — live data could not be retrieved.")
}
```

## Row {height="18%"}

::: {.card title="Why this matters for West Boundary agriculture"}
Summer is when agricultural water demand peaks and natural supply is lowest. The
West Boundary has faced repeated drought stress, including the **2021 Temporary
Protection Order** in the West Kettle watershed that restricted withdrawals.

The envelope chart and the comparisons above put **today's flows in the context
of decades of record** (48–70 years per main station), so producers and planners
can quickly see whether the watershed is tracking toward a low-flow year. When
the red line sits near or below the lower edge of the blue band, the watershed is
running drier than usual for the season — an early signal for storage,
conservation, and engagement decisions.
:::

## Row {height="55%"}

### Column {width="55%"}

```{r}
#| title: "Long-term trend — summer low flow (7-day minimum, Jul–Sep)"

sl <- flow_summer %>%
  mutate(fit = predict(lm(summer_7day_min ~ year, data = flow_summer)))

plot_ly(sl, x = ~year, y = ~summer_7day_min, type = "scatter", mode = "markers",
        marker = list(color = "#1f77b4", size = 7),
        hovertemplate = "Year %{x}: %{y:.1f} m³/s<extra></extra>", showlegend = FALSE) %>%
  add_lines(x = ~year, y = ~fit, line = list(color = "#d62728", dash = "dash"),
            name = "Linear fit", inherit = TRUE) %>%
  layout(
    yaxis = list(title = "7-day minimum flow (m³/s)"),
    xaxis = list(title = ""),
    annotations = list(
      x = 0.02, y = 0.98, xref = "paper", yref = "paper", xanchor = "left",
      text = paste0(
        "Mann-Kendall: ", trend_row$direction[1],
        " (τ = ", trend_row$tau[1],
        ", p = ", trend_row$p_value[1], ")<br>",
        "Sen's slope ≈ ", trend_row$sen_slope_m3s_yr[1], " m³/s per year"
      ),
      showarrow = FALSE, align = "left",
      font = list(size = 11), bgcolor = "rgba(255,255,255,0.85)"
    )
  )
```

### Column {width="45%"}

::: {.card title="What the trend means"}
The **7-day summer low flow** is the metric regulators and drought planners watch
most closely — it captures the sustained low-water conditions that stress irrigation
and aquatic habitat.

A **declining long-term trend** in summer lows is consistent with warmer summers,
earlier snowmelt, and increasing water demand in the West Boundary. It does not
predict any single year, but it frames why storage, efficiency, and cooperative
planning matter even in non-drought years.
:::

# Snowpack & Water Supply

## Row {height="22%"}

```{r}
#| content: valuebox
#| title: "Grano Creek SWE (2E07P)"
list(
  icon  = "snow",
  color = snow_vb,
  value = if (is.na(snow_swe)) "n/a" else paste0(snow_swe, " mm")
)
```

```{r}
#| content: valuebox
#| title: "SWE vs. normal for today"
list(
  icon  = "thermometer-snow",
  color = snow_vb,
  value = if (is.na(snow_pct)) "n/a" else paste0(snow_pct, "% of median")
)
```

```{r}
#| content: valuebox
#| title: "Snowpack status"
list(
  icon  = "cloud-snow",
  color = snow_vb,
  value = snow_status
)
```

```{r}
#| content: valuebox
#| title: "Station elevation"
list(
  icon  = "arrow-up",
  color = "secondary",
  value = paste0(snow_station$elevation_m[1], " m")
)
```

## Row {height="78%"}

### Column {width="62%"}

```{r}
#| title: "Snow water equivalent vs. historical normal — Grano Creek (leading indicator for summer supply)"

wy_start <- as.Date(paste0(snow_wy - 1L, "-10-01"))
senv <- snow_normals %>%
  filter(wyday <= 366) %>%
  arrange(wyday) %>%
  mutate(date = as.Date(wyday - 1, origin = wy_start))

scur <- snow_line %>% mutate(date = Date)

plot_ly() %>%
  add_ribbons(data = senv, x = ~date, ymin = ~p10, ymax = ~p90,
              name = "Historical SWE range (10–90%)", line = list(width = 0),
              fillcolor = "rgba(23,190,207,0.15)", hoverinfo = "skip") %>%
  add_ribbons(data = senv, x = ~date, ymin = ~p25, ymax = ~p75,
              name = "Typical SWE range (25–75%)", line = list(width = 0),
              fillcolor = "rgba(23,190,207,0.30)", hoverinfo = "skip") %>%
  add_lines(data = senv, x = ~date, y = ~p50, name = "Median SWE",
            line = list(color = "#17becf", dash = "dash", width = 2)) %>%
  add_lines(data = scur, x = ~date, y = ~swe,
            name = paste0("WY ", snow_wy, " observed"),
            line = list(color = "#9467bd", width = 3)) %>%
  layout(
    yaxis = list(title = "Snow water equivalent (mm)"),
    xaxis = list(title = "", tickformat = "%b"),
    hovermode = "x unified",
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.08)
  )
```

### Column {width="38%"}

::: {.card title="Why snowpack matters here"}
**Grano Creek (2E07P)** is the Boundary Region's only **real-time automated snow
station** in the Kettle River watershed. Snow that accumulates at elevation is the
reservoir that feeds spring freshet and summer baseflow — low or early-melting
snowpack is a **leading indicator** that summer irrigation supply may be tight,
often months before streamflow hits critical levels.

The RDKB drought updates and provincial snow bulletins use this station to track
whether the watershed entered the growing season with adequate mountain storage.
:::

```{r}
#| title: "Snow station"

data.frame(
  Station = snow_station$station_name,
  ID = snow_station$station_id,
  Elevation = paste0(snow_station$elevation_m, " m"),
  Record = paste0(snow_station$start_year, "–", snow_station$end_year)
) %>% kable()
```

# Groundwater & Aquifers

## Row {height="22%"}

```{r}
#| content: valuebox
#| title: "Grand Forks well (OW217) — depth to water"
gw_box("OW217")
```

```{r}
#| content: valuebox
#| title: "Beaverdell well (OW306) — depth to water"
gw_box("OW306")
```

```{r}
#| content: valuebox
#| title: "Midway well (OW444) — depth to water"
gw_box("OW444")
```

```{r}
#| content: valuebox
#| title: "Observation wells below normal"
list(
  icon  = "reception-2",
  color = if (gw_below > 0) "warning" else "success",
  value = if (gw_n == 0) "n/a" else paste0(gw_below, " of ", gw_n)
)
```

## Row {height="78%"}

### Column {width="62%"}

```{r}
#| title: "Water-table depth vs. the historical normal range — Grand Forks (OW217). Lower on the chart = deeper water table = drier."

gw_hero <- "OW217"
genv <- gw_normals %>%
  filter(well == gw_hero, yday <= 365) %>%
  arrange(yday) %>%
  mutate(date = as.Date(yday - 1, origin = paste0(yr, "-01-01")))

gcur <- if (have_gw) {
  gw_live %>%
    filter(well == gw_hero, Date >= as.Date(paste0(yr, "-01-01"))) %>%
    arrange(Date)
} else {
  data.frame(Date = as.Date(character()), depth = numeric())
}

plot_ly() %>%
  add_ribbons(data = genv, x = ~date, ymin = ~p10, ymax = ~p90,
              name = "Historical range (10–90%)", line = list(width = 0),
              fillcolor = "rgba(44,160,44,0.12)", hoverinfo = "skip") %>%
  add_ribbons(data = genv, x = ~date, ymin = ~p25, ymax = ~p75,
              name = "Typical range (25–75%)", line = list(width = 0),
              fillcolor = "rgba(44,160,44,0.28)", hoverinfo = "skip") %>%
  add_lines(data = genv, x = ~date, y = ~p50, name = "Median (normal)",
            line = list(color = "#2ca02c", dash = "dash", width = 2)) %>%
  add_lines(data = gcur, x = ~Date, y = ~depth, name = paste(yr, "observed"),
            line = list(color = "#8c564b", width = 3)) %>%
  layout(
    yaxis = list(title = "Depth to water below ground (m)", autorange = "reversed"),
    xaxis = list(title = "", tickformat = "%b"),
    hovermode = "x unified",
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.08)
  )
```

### Column {width="38%"}

```{r}
#| title: "Observation wells (B.C. PGOWN)"

gw_current %>%
  transmute(
    Well = well,
    Location = location,
    Aquifer = aquifer,
    `Depth (m)` = round(depth, 2),
    Record = ifelse(is.na(start_year), "—", paste0(start_year, "–", end_year)),
    Status = ifelse(is.na(status), "no baseline", status)
  ) %>%
  kable()
```

::: {.card title="Why groundwater matters here"}
The Kettle valley aquifers are **strongly connected to the river** — the Grand
Forks aquifer sits next to the Kettle/Granby confluence and rises and falls with
the freshet. When both streamflow **and** water tables are running below normal at
the same time, it points to a basin-wide moisture deficit rather than a local
blip — an important signal for irrigation timing, storage, and shared
conservation decisions.
:::

# Stations & Live Data

## Row {height="40%"}

```{r}
#| title: "Monitoring stations (Water Survey of Canada)"

stations_meta %>%
  arrange(STATION_NUMBER) %>%
  transmute(
    `Station ID` = STATION_NUMBER,
    Name = STATION_NAME,
    `Drainage area (km²)` = round(DRAINAGE_AREA_GROSS),
    `Record` = ifelse(is.na(start_year), "—", paste0(start_year, "–", end_year)),
    `Years` = n_years
  ) %>%
  kable()
```

## Row {height="60%"}

```{r}
#| title: "Daily mean discharge — all monitored sites, current year"

if (have_live) {
  lab <- setNames(stations_meta$STATION_NAME, stations_meta$STATION_NUMBER)
  live2 <- live %>% mutate(name = lab[STATION_NUMBER])
  plot_ly(live2, x = ~Date, y = ~flow, color = ~name, type = "scatter",
          mode = "lines") %>%
    layout(yaxis = list(title = "Discharge (m³/s)"), xaxis = list(title = ""),
           hovermode = "x unified",
           legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.12))
} else {
  cat("Live data could not be retrieved at build time.")
}
```

# Agriculture & Field Observations

## Row {height="22%"}

```{r}
#| content: valuebox
#| title: "Irrigation season"
list(
  icon  = "moisture",
  color = if (irrigation_active) "primary" else "secondary",
  value = if (irrigation_active) "Active" else "Off"
)
```

```{r}
#| content: valuebox
#| title: "Flow vs. demand"
list(
  icon  = "droplet",
  color = vb_color[[hstatus]],
  value = if (irrigation_active) hstatus else "Off"
)
```

## Row {height="78%"}

```{r}
#| title: "Streamflow through the irrigation window (Apr–Oct) — Kettle River near Westbridge"

irr_start <- as.Date(paste0(yr, "-04-01"))
irr_end   <- as.Date(paste0(yr, "-10-31"))

plot_ly() %>%
  add_ribbons(
    data = env %>% filter(date >= irr_start, date <= irr_end),
    x = ~date, ymin = ~p25, ymax = ~p75,
    name = "Typical range (irrigation months)", line = list(width = 0),
    fillcolor = "rgba(44,160,44,0.20)", hoverinfo = "skip"
  ) %>%
  add_lines(
    data = env %>% filter(date >= irr_start, date <= irr_end),
    x = ~date, y = ~p50, name = "Median",
    line = list(color = "#2ca02c", dash = "dash", width = 2)
  ) %>%
  add_lines(
    data = cur_line %>% filter(Date >= irr_start, Date <= irr_end),
    x = ~Date, y = ~flow, name = paste(yr, "observed"),
    line = list(color = "#d62728", width = 3)
  ) %>%
  layout(
    yaxis = list(title = "Daily mean discharge (m³/s)"),
    xaxis = list(title = "", tickformat = "%b"),
    hovermode = "x unified",
    legend = list(orientation = "h", x = 0.5, xanchor = "center", y = 1.08)
  )
```

# About & Data {#about-data}

## Row

### Column

::: {.card title="What this dashboard is"}
A **pilot dashboard** for the RDKB West Boundary Agricultural Water Supply and
Storage planning process. It brings public hydrometric information together in an
accessible place to support **planning and community engagement** — helping
producers, First Nations, local government, and agencies share a common picture
of seasonal water conditions and drought risk in the Kettle River watershed from
Midway upstream.

It is **not** a regulatory monitoring tool. Readings are provisional and
unverified, and should not be used for compliance or enforcement.
:::

::: {.card title="Data sources & methods"}
- **Live streamflow:** Environment and Climate Change Canada (ECCC) real-time
  hydrometric web service, retrieved with the open-source
  [`tidyhydat`](https://docs.ropensci.org/tidyhydat/) R package (parameter 47,
  daily mean, m³/s).
- **Streamflow normals:** Water Survey of Canada **HYDAT** database, precomputed by
  `scripts/build_flow_normals.R` and `scripts/build_flow_analytics.R` (MAD
  thresholds, drought analog years, summer low-flow trend).
- **Groundwater:** B.C. Provincial Groundwater Observation Well Network (PGOWN),
  precomputed by `scripts/build_gw_normals.R`.
- **Snowpack:** B.C. Automated Snow Weather Station **Grano Creek (2E07P)** via
  the B.C. Data Catalogue (`scripts/build_snow_normals.R`).
- **Status classes:** *Severe Low* (least water available — below the 10th
  percentile for the day), then *Low*, *Normal*, *Above Normal*, *High*.
:::

### Column

::: {.card title="Confidentiality & responsible use"}
This dashboard uses **public and voluntarily shared information only**. Should
producer-contributed inputs (for example well levels, repeat photo points, or
irrigation timing) be added in future, they would be **aggregated or anonymised**
so that individual operations cannot be identified, consistent with the project's
commitment to a transparent, non-regulatory process and with FOIPPA.
:::

::: {.card title="Ownership, hosting & upkeep"}
Built entirely with **open-source tools** (Quarto + R) so it can be fully owned,
hosted, and maintained by the Regional District. All source files, data-build
scripts, and documentation are included in the project repository. Refreshing the
historical baselines is a single command (`Rscript scripts/build_flow_normals.R`);
live readings update automatically each time the site is rebuilt.

*Potential future inputs identified in the RFP — groundwater well levels,
date-stamped/repeat photo points, and irrigation season timing — can be added as
additional pages using the same pattern.*
:::
