---
title: "West Boundary Water Security"
author: "TPB Economics Team"
format:
  html:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

# Overview

Near-real-time surface-water monitoring for the West Boundary region. Daily mean
discharge is pulled directly from the Environment and Climate Change Canada (ECCC)
real-time hydrometric web service via the [`tidyhydat`](https://docs.ropensci.org/tidyhydat/)
R package.

```{r}
#| include: false

# 1. LOAD REQUIRED PACKAGES
library(tidyverse)
library(plotly)
library(tidyhydat)

# 2. AUTOMATED TELEMETRY INGESTION (SURFACE WATER)
target_stations <- c("08NN003", "08NN026")

# Fetch live data from ECCC hydrometric web service. The fetch is wrapped in
# tryCatch so a temporary network/service outage never breaks the site build.
live_flow_data <- tryCatch(
  realtime_ws(
    station_number = target_stations,
    parameters = 47,
    start_date = Sys.Date() - 60,
    end_date = Sys.Date()
  ) %>%
    mutate(Date = as.Date(Date)) %>%
    group_by(STATION_NUMBER, Date) %>%
    summarize(Daily_Mean_Flow = mean(Value, na.rm = TRUE), .groups = "drop"),
  error = function(e) {
    message("Live hydrometric fetch failed: ", conditionMessage(e))
    tibble(
      STATION_NUMBER = character(),
      Date = as.Date(character()),
      Daily_Mean_Flow = numeric()
    )
  }
)

have_data <- nrow(live_flow_data) > 0

kettle_flow   <- live_flow_data %>% filter(STATION_NUMBER == "08NN003")
boundary_flow <- live_flow_data %>% filter(STATION_NUMBER == "08NN026")

current_kettle_flow <- if (nrow(kettle_flow) > 0) {
  kettle_flow %>% filter(Date == max(Date)) %>% pull(Daily_Mean_Flow) %>% round(1)
} else {
  NA_real_
}
```

# Kettle River — Last 60 Days Live Telemetry

```{r}
#| echo: false

if (have_data) {
  plot_ly(kettle_flow, x = ~Date) %>%
    add_lines(
      y = ~Daily_Mean_Flow,
      name = "Kettle River (Ferry)",
      line = list(color = "#1f77b4", width = 3)
    ) %>%
    add_lines(
      data = boundary_flow,
      y = ~Daily_Mean_Flow,
      name = "Boundary Creek (Midway)",
      line = list(color = "#7f7f7f", width = 2, dash = "dot")
    ) %>%
    layout(
      xaxis = list(title = ""),
      yaxis = list(title = "Discharge (m³/s)"),
      hovermode = "x unified",
      legend = list(orientation = "h", x = 0.5, xanchor = "center", y = -0.1)
    )
} else {
  cat(
    "> Live hydrometric data is temporarily unavailable.",
    "The chart will populate automatically the next time the site is rebuilt",
    "with the data service reachable."
  )
}
```

# Methodology

+ **Automated Hydrology:** Current flow data is automatically retrieved from the
  ECCC real-time hydrometric API via the `tidyhydat` R package (parameter 47,
  discharge, in m³/s).

+ **Spatial Data Integrity:** Land classification (ALR parcels) can be queried
  directly from the BC Data Catalogue API using the `bcdata` R package, ensuring
  models always use the most authoritative and up-to-date regional boundaries.
