  ---
title: "West Boundary Water Security"
author: "Your Name/Portfolio"
format:
dashboard:
theme: yeti
orientation: rows
execute:
echo: false
warning: false
message: false
---

```{r}
#| include: false
#| echo: false
# 1. LOAD REQUIRED PACKAGES
# Ensure you have installed these via: install.packages(c("tidyhydat", "bcdata", "sf", "bcmaps", "tidyverse", "plotly", "leaflet", "bslib"))
library(tidyverse)
library(plotly)
library(leaflet)
library(bslib)
library(tidyhydat) 
library(bcdata)    
library(sf)

# 2. AUTOMATED TELEMETRY INGESTION (SURFACE WATER)
target_stations <- c("08NN003", "08NN026")

# Fetch live data from ECCC hydrometric web service
live_flow_data <- realtime_ws(
  station_number = target_stations,
  parameters = 47, 
  start_date = Sys.Date() - 60,
  end_date = Sys.Date()
) %>%
  mutate(Date = as.Date(Date)) %>%
  group_by(STATION_NUMBER, Date) %>%
  summarize(Daily_Mean_Flow = mean(Value, na.rm = TRUE), .groups = "drop")

kettle_flow <- live_flow_data %>% filter(STATION_NUMBER == "08NN003")
boundary_flow <- live_flow_data %>% filter(STATION_NUMBER == "08NN026")

current_kettle_flow <- kettle_flow %>% filter(Date == max(Date)) %>% pull(Daily_Mean_Flow) %>% round(1)

# 3. SPATIAL DATA INGESTION
# Query ALR polygons from BC Data Catalogue
# alr_query <- bcdc_query_geodata("alc-agricultural-land-reserve-polygons") %>%
#   filter(BBOX(c(-119.0, 49.0, -118.7, 49.2))) %>% 
#   collect()
```



<!-- # Real-Time Telemetry -->
<!--  ```{r telemetry} -->

<!-- #| content: valuebox -->
<!-- #| title: "Current Flow: Kettle River (08NN013)" -->
<!-- list( -->
<!--   icon = "droplet-half", -->
<!--   color = ifelse(current_kettle_flow < 20, "danger", "primary"), -->
<!--   value = paste0(current_kettle_flow, " m³/s") -->
<!-- ) -->


<!-- #| content: valuebox -->
<!-- #| title: "Live Data Source" -->
<!-- list( -->
<!--   icon = "cloud-arrow-down", -->
<!--   color = "secondary", -->
<!--   value = "ECCC Web Service" -->
<!-- ) -->
<!-- ``` -->



# Kettle River - Last 30 Days Live Telemetry
```{r last 30}
#| echo: false

plot_ly(kettle_flow, x = ~Date) %>%
  add_lines(y = ~Daily_Mean_Flow, name = "Kettle River (Ferry)", line = list(color = "#1f77b4", width = 3)) %>%
  add_lines(data = boundary_flow, y = ~Daily_Mean_Flow, name = "Boundary Creek (Midway)", line = list(color = "#7f7f7f", width = 2, dash = "dot")) %>%
  layout(
    xaxis = list(title = ""),
    yaxis = list(title = "Discharge (m³/s)"),
    hovermode = "x unified",
    legend = list(orientation = 'h', x = 0.5, xanchor = 'center', y = -0.1)
  )
```


<!-- # Spatial Planning -->

<!-- ```{r spatial planning} -->

<!-- leaflet() %>% -->
<!--   addProviderTiles(providers$CartoDB.Positron) %>% -->
<!--   setView(lng = -118.85, lat = 49.05, zoom = 11) %>% -->
<!--   addPolygons(data = alr_query,  -->
<!--               color = "#8c564b",  -->
<!--               weight = 1,  -->
<!--               fillOpacity = 0.2,  -->
<!--               popup = ~paste("ALR Parcel ID:", ALC_DOCUMENT_ID)) %>% -->
<!--   addLegend("bottomright",  -->
<!--             colors = c("#8c564b"),  -->
<!--             labels = c("ALR Boundaries (bcdata)"),  -->
<!--             title = "Land Base") -->
<!-- ``` -->

# Methodology

+ Automated Hydrology: Current flow data is automatically retrieved from the ECCC real-time hydrometric API via the tidyhydat R package.

+ Spatial Data Integrity: Land classification (ALR parcels) is queried directly from the BC Data Catalogue API using the bcdata R package, ensuring models always use the most authoritative and up-to-date regional boundaries.